Report on the Fourth Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE4)

### 2017-03-08 WSSSPE4_report/WSSSPE4_report.tex
added a few more lightning talk descriptions

Intro and description.
A \textsuperscript{\textasteriskcentered} indicates which author presented the
lightning talk, for those with multiple authors.
%
\begin{itemize}
    \item \textbf{Shoaib Sufi: \emph{The Software Sustainability Institute Fellowship
    programme; supporting the social side of research software}}~\cite{Sufi:2016ws}.
    Sufi, Community Lead at the Software Sustainability Institute (SSI), discussed
    the SSI Fellowship program and the fellows, including their
    demographic breakdown and domains of practice. This program came about
    based on the institute's desire to integrate with the research community---thus
    inspiring change from within---rather than preach to it. Five SSI fellows
    attended WSSSPE4.

    \item \textbf{Sandra Gesing, Maytal Dahan, Linda B.~Hayden, Katherine Lawrence,
    Suresh Marru, Marlon Pierce, Nancy Wilkins-Diehr, and Michael Zentner:
    \emph{The Science Gateways Community Institute - Supporting Communities to
    Achieve Sustainability for Their Science Gateways}}~\cite{Gesing:2016ws}.
    \note{Description}

    \item \textbf{Karthik Ram\textsuperscript{\textasteriskcentered}, Noam Ross,
    and Scott Chamberlain: \emph{A model for
    peer review and onboarding research software}}~\cite{Ram:2016ws}.
    \note{Description}
    %Some people greatly encouraged to only submit to a journal code that has been peer reviewed

    \item \textbf{Frank L\"{o}ffler and Steven Brandt\textsuperscript{\textasteriskcentered}:
    \emph{A vision of computing in 10+ years}}~\cite{Loffler:2016ws}.
    \note{Description}
    %“We (need to) decouple science from numerical methods and optimization” Hard open problem.

>>>>

Intro and description.
A \textsuperscript{\textasteriskcentered} indicates which author presented the
lightning talk, for those with multiple authors.
%
\begin{itemize}[itemsep=1ex]
    \item \textbf{Shoaib Sufi: \emph{The Software Sustainability Institute Fellowship
    programme; supporting the social side of research software}}~\cite{Sufi:2016ws}.
    Sufi, Community Lead at the Software Sustainability Institute (SSI), discussed
    the SSI Fellowship program and the fellows, including their
    demographic breakdown and domains of practice. This program came about
    based on the institute's desire to integrate with the research community---thus
    inspiring change from within---rather than preach to it. Five SSI fellows
    attended WSSSPE4.

    \item \textbf{Sandra Gesing, Maytal Dahan, Linda B.~Hayden, Katherine Lawrence,
    Suresh Marru, Marlon Pierce, Nancy Wilkins-Diehr, and Michael Zentner:
    \emph{The Science Gateways Community Institute - Supporting Communities to
    Achieve Sustainability for Their Science Gateways}}~\cite{Gesing:2016ws}.
    In this talk, Gesing introduced the concept of science gateways: end-to-end software
    solutions for a community that hide underlying complexity. She also posed the need
    for gateways, which include the increased complexity of research and associated
    hardware\slash software, a greater need for reproducibility and openness, and the
    opportunity to integrate research with teaching for better workforce development.
    In fact, gateway users surpassed login-based users in XSEDE systems in 2013,
    and 1.4 times as many users accessed XSEDE via gateways compared with
    login in March 2016. The Science Gateways Community Institute offers support
    related to gateways in five areas: Incubator, Extended Developer Support, Scientific
    Software Collaborative, Community Engagement and Exchange, and Workforce
    Development.

    \item \textbf{Karthik Ram\textsuperscript{\textasteriskcentered}, Noam Ross,
    and Scott Chamberlain: \emph{A model for
    peer review and onboarding research software}}~\cite{Ram:2016ws}.
    This talk described the rOpenSci community's review process for research
    software packages. Submission of a package starts with a pre-submission inquiry
    for fit with the community criteria, then peer reviewers evaluate based on
    usability, quality, and style. Upon acceptance, packages are given an rOpenSci
    README badge and added to the system. Important components include a code of
    conduct, open and non-adversarial review, and that no submission is
    rejected---just out of scope. Remaining challenges include scaling, author
    incentives, and automation of the process.

    \item \textbf{Frank L\"{o}ffler and Steven Brandt\textsuperscript{\textasteriskcentered}:
    \emph{A vision of computing in 10+ years}}~\cite{Loffler:2016ws}.
    This talk began by describing the challenges facing science, and particularly
    computational science, in recent years: the amount and complexity of knowledge
    has increased dramatically, making it difficult for individuals or even teams to
    manage modern, complex problems.
    Brandt submitted an argument for decoupling science simulation and
    the solution of computational problems away from numerical methods and low-level
    optimization. This may be possible with domain-specific languages, or the use
    of modern features in existing languages. This vision also includes reuse of
    basic computing infrastructure, including common data structures, parameter files,
    and high-performance I\slash O. This computing future necessarily requires
    better recognition and reward systems for research software developers (i.e.,
    research software engineers), and ways to make increasingly complex scientific
    papers more understandable.
